Basic regression in Gaussian processes

### 2019-10-29 notebook/gp-manual-vs-greta.Rmd
gp greta exploration

---
layout: post
title: Basic regression in Gaussian processes 
category: ecology
tag: nonparametric-bayes
code: true
output: html_document
---

@@

and plot our solution (mean, 2 standard deviations, and the ranom samples.)

@@

ggplot(dat,aes(x=x,y=value)) + 
  #geom_line(aes(group=rep), color =  rgb(0.7, 0.1, 0.4), alpha = 0.2) + #REPLICATES
  geom_ribbon(data = gp, 
              aes(x, 
                  y = Ef, 
                  ymin = Ef - sigma, 
                  ymax = Ef + sigma),
              fill="blue", alpha = 0.1) +
  geom_line(dat = gp, aes(x=x,y=Ef), size=1) + #MEAN
  geom_point(data=obs,aes(x=x,y=y)) +  #OBSERVED DATA
  scale_y_continuous(lim=c(-3,3), name="output, f(x)") +
  xlab("input, x")
```

@@

and plot

```{r}
dat <- data.frame(x=x_predict, t(values))
dat <- melt(dat, id="x")

@@

>>>>

---
layout: post
title: Basic regression in Gaussian processes 
category: ecology
tag: nonparametric-bayes
code: true
output: github_document
---

@@

and plot our solution (mean in black, 2 standard deviations in grey riboon, and the 200 random samples in purple.)

@@

ggplot(dat,aes(x=x,y=value)) + 
  geom_line(aes(group=rep), color =  rgb(0.7, 0.1, 0.4), alpha = 0.2) + #REPLICATES
  geom_ribbon(data = gp, 
              aes(x, 
                  y = Ef, 
                  ymin = Ef - sigma, 
                  ymax = Ef + sigma),
              fill="grey", alpha = 0.4) +
  geom_line(dat = gp, aes(x=x,y=Ef), size=1) + #MEAN
  geom_point(data=obs,aes(x=x,y=y)) +  #OBSERVED DATA
  scale_y_continuous(lim=c(-3,3), name="output, f(x)") +
  xlab("input, x")
```

@@

and plot (both sample curves, in purple, and the theoretical expected variance as a grey ribbon)

```{r}
dat <- data.frame(x=x_predict, t(values))
dat <- reshape2::melt(dat, id="x")

@@

# Try greta with fixed parameters

Here I try to replicate the above results under 'no additive noise' by setting the noise term to (almost) zero.  (trying `obs_sd = 0` causes the MCMC to fail to find initial values)

# hyperparameters
rbf_var <- 1
rbf_len <- 1
obs_sd <- 0.001

```{r}
# fit the model by Hamiltonian Monte Carlo
m <- model(f_plot)
draws <- mcmc(m, chains = 1, one_by_one = TRUE)
```

Unfortunately, instead of causing the GP to collapse down only at the data points, like in the example above, this cause the GP samples to collapse entirely:

```{r}
draws[[1]] %>% t() %>% as_tibble() %>% dplyr::select(1:200) %>% 
  mutate(x = x_predict) %>% pivot_longer(-x)  %>%
  ggplot(aes(x,value)) + 
  geom_line(aes(group=name), alpha=0.3, col=rgb(0.7, 0.1, 0.4, 0.1)) +
  geom_point(data = obs, aes(x,y))
```

# MCMC with greta

### 2019-10-29 notebook/greta-gp.Rmd
gp greta exploration

```{r setup}
library(tidyverse)
library(greta)
library(greta.gp)
#tensorflow::use_session_with_seed(1234)

@@

```{r}
ghost <- readr::read_csv("../data/ghost_sims.csv.gz")
wide <- ghost %>% pivot_wider(names_from = reps, values_from = x) %>% select(-t)
n <- dim(wide)[1]
x_t1 <- wide[-1,]
x_t <- wide[-n,] 
y <- x_t1 - x_t
Y <- y %>% apply(1, mean, na.rm=TRUE)
X <- x_t %>% apply(1, mean, na.rm=TRUE)

@@

```{r}
# fit the model by Hamiltonian Monte Carlo
m <- model(f_plot, rbf_len, rbf_var, obs_sd)
draws <- mcmc(m, chains = 1, one_by_one = TRUE)
```

>>>>

```{r setup}
library(greta)
library(greta.gp)
#tensorflow::use_session_with_seed(1234)
library(tidyverse)

@@

```{r eval=FALSE}
ghost <- readr::read_csv("../data/ghost_sims.csv.gz")
wide <- ghost %>% dplyr::select(-t) %>% pivot_wider(names_from = reps, values_from = x) 
n <- dim(wide)[1]
x_t1 <- wide[-1,]
x_t <- wide[-n,] 
y <- x_t1 - x_t
Y <- y %>% apply(1, mean, na.rm=TRUE)
X <- x_t %>% apply(1, mean, na.rm=TRUE)

@@

```{r}
# fit the model by Hamiltonian Monte Carlo
m <- model(f, rbf_len, rbf_var, obs_sd)
draws <- mcmc(m, chains = 1, one_by_one = TRUE)
```
